What's hot and what's not:\\Windowed developer topic analysis

### 2009-04-11 lda-paper.tex
some more edits but it puts it at 11 pages ;_;

  As development on a software project progresses, developers will
  naturally shift their focus between different topics and tasks many
  times.  Managers and newcomer developers often seek ways of
  understanding what tasks have recently been worked on and how much
  effort has gone into each; for example, a manager might wonder what
  unexpected tasks occupied their team's attention during a period
  when they were supposed to have been implementing a set of new
  features.  Tools such as Latent Dirichlet Allocation (LDA) and
  Latent Semantic Indexing (LSI) can be used to analyze check-in
  comments over the entire lifetime of a project. Previous work on
  developer topic analysis has leveraged these tools to associate
  commit log comments with independent topics extracted from these
  commit log comments.  In this work, we use LDA and LSI to analyze
  periods, such as months, within a project's lifetime to create a
  time-windowed model of changing development topics.  We propose a
  visualization of this model that allows us to explore the evolving
  stream of topics of development occurring over time.  We demonstrate
  that windowed topic analysis offers advantages over topic analysis
  applied to a projects lifetime because many topics are quite local.

@@

Commits in SCSs can have
multiple purposes. Thus we hope that our research will help uncover
these multiple purposes that manifest themselves as topics interleaved
within the commits.  Others have demonstrated that 
topic analysis techniques, such as Latent Semantic Indexing (LSI) and
Latent Dirichlet Allocation (LDA), can aid in separating topics from
these mixed comments. These topics often describe distinct ideas that
are interleaved throughout the development. A topic represents a
groups of commit log comments that are related to each other by their
content.  In this case a topic is a set of tokens.

@@

We propose to extend existing topic analysis techniques by
applying them to windows of documents over time. We hope that by
looking %XX X ADDED
at topic clusters during periods or windows over time we can identify
ongoing areas of focus on a select group of topics as well reflect more intermittent topics. Figure \ref{fig:lda} illustrates
the kind of visualization we want to automatically generate. We want
to see topics that are unique to a certain time period, displayed
across the time axis. Those topics that recur, or occur over a
larger period are plotted continuously. In our example we have
titled each topic with a single word.

In this paper we explore how topics shift over time in the source
control system (SCS) of a project, using several open source database
systems as examples. We analyze commit log comments in each
time window and we extract the topics of that time window. We expect
that topics will change over time and the similarities and differences
in these topics will indicate developer focus and changes in developer
focus as it changes over time.

@@

We now define some terms that we will use in the rest of the paper: A
\emph{message} is a block of text written by developers. In this
paper, messages will be the CVS commit log comments made when the user
commits changes to files in a repository. A \emph{word distribution}
is the summary of a message by word count. Each word distribution will
be over the words in all messages. However, most words will not appear
in a message, a word distribution is effectively a word count divided
by the message size. A \emph{topic} is a word distribution, a
set of words that form a word distribution that is unique and
independent within the set of documents in our total corpus. One could
think of a topic as a distribution of the centroid of a group of
messages. In this paper we often summarize topics by the top 10 most
frequent words of their word distribution.  A \emph{trend} is a one or
more similar topics that recur over time.  Trends are particularly
important, as they indicate long-lived and recurring topics that may
provide key insights into the development of the project.

@@

A sampling of the notable topic words is displayed in Table
\ref{tab:portability}, we chose topics that we felt confident we could
name.  To name each topic we selected a term that seemed to best
summarized that topic.  After extracting these topics, we attempted to
track the evolution of topics by visualizing the topics and joining
similar topics into trends.  Figure \ref{fig:lda} displays a manually
created sample plot of the extracted topics in Table
\ref{tab:portability}.

@@

Our methodology is to first extract the commits from SCS repositories
such as CVS or Bitkeeper. Then we extract the revisions and commits,
grouping the revisions into commits if necessary. We extract the
commit log comment of each commit, count the occurrence of each word
in the message, remove stop words, and then produce word distributions
for each message.

@@

We mirrored the repositories and their revisions using software such
as rsync, CVSSuck, \texttt{softChange} and bt2csv.
\texttt{softChange} provided a general schema for storing
revisions. CVSSuck and rsync mirrored CVS repositories while bt2csv
mirrored web accessible BitKeeper repositories.

The documents we are analyzing are the commit log comments i.e., the
comments that are added when revisions to the project are committed.
Documents are converted into word count distributions. These
distributions are then normalized by the size of the message. After
all messages are processed the distributions are extended to include
all words from all of the  distributions.

@@

Given all messages, we group the messages into windows. We could use
overlapping windows, but in this paper we use non-overlapping windows
of a set time unit because it simplifies analysis.  Windowing by time
allows for many levels of granularity.  We usually used one month as
the length of our time windows. While we could use different window
lengths for the study we think that a month is a sizable unit of
development, which is large enough to show shifts in focus, but coarse
enough not to show too much detail.

>>>>

  As development on a software project progresses, developers will
  naturally shift their focus between different topics and tasks many
  times.  Managers and newcomer developers often seek ways of
  understanding what tasks have recently been worked on and how much
  effort has gone into each; for example, a manager might wonder what
  unexpected tasks occupied their team's attention during a period
  when they were supposed to have been implementing a set of new
  features.  Tools such as Latent Dirichlet Allocation (LDA) and
  Latent Semantic Indexing (LSI) can be used to analyze check-in
  comments over the entire lifetime of a project. Previous work on
  developer topic analysis has leveraged these tools to associate
  commit log comments with independent topics extracted from these
  commit log comments.  In this work, we use LDA  to analyze
  periods, such as months, within a project's lifetime to create a
  time-windowed model of changing development topics.  We propose 
  visualizations of this model that allows us to explore the evolving
  stream of topics of development occurring over time.  We demonstrate
  that windowed topic analysis offers advantages over topic analysis
  applied to a projects lifetime because many topics are quite local.

@@

%XXX RM 
Commits in source control systems (SCSs) can have multiple
purposes. Thus we hope that our research will help uncover these
multiple purposes that manifest themselves as topics interleaved
within the commits.  Others have demonstrated that topic analysis
techniques, such as Latent Semantic Indexing (LSI) and Latent
Dirichlet Allocation (LDA), can aid in separating topics from these
mixed comments. These topics often describe distinct ideas that are
interleaved throughout the development. A topic represents both a word
distribution and a group of commit log comments that are related to
each other by their content.  In this case a topic is a set of tokens.

@@

We propose to extend existing topic analysis techniques by applying
them to windows of documents over time. We hope that by
looking %XX X ADDED
at topic clusters during periods or windows over time we can identify
ongoing areas of focus on a select group of topics as well reflect
more intermittent topics. Figure \ref{fig:lda} illustrates the kind of
visualization we want to automatically generate. We want to see topics
that are unique to a certain time period, displayed across the time
axis. Those topics that recur, or occur over a larger period are
plotted continuously. In our example we have titled each topic with a
single word drawn from itself.

In this paper we explore how topics shift over time in the SCS of a
project, using several open source database systems as examples. We
analyze commit log comments in each time window and we extract the
topics of that time window. We expect that topics will change over
time and the similarities and differences in these topics will
indicate developer focus and changes in developer focus as it changes
over time.

@@

We now define some terms that we will use in the rest of the paper: A
\emph{message} is a block of text written by developers. In this
paper, messages will be the CVS and BitKeeper commit log comments made
when the user commits changes to files in a repository. A \emph{word
  distribution} is the summary of a message by word count. Each word
distribution will be over the words in all messages. However, most
words will not appear in a message, a word distribution is effectively
a word count divided by the message size. A \emph{topic} is a word
distribution, a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top 10 most frequent words of their word distribution.  A \emph{trend}
is a one or more similar topics that recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project.

@@

A sampling of the notable topic words is displayed in Table
\ref{tab:portability}, we chose topics that we felt confident we could
name.  To name each topic we selected a term that seemed to best
summarize that topic.  After extracting these topics, we attempted to
track the evolution of topics by visualizing the topics and joining
similar topics into trends.  Figure \ref{fig:lda} displays a manually
created sample plot of the extracted topics in Table
\ref{tab:portability}.

@@

%XXX Smaller?

Our methodology is to first extract the commit log comments from a
project's SCS repository. We filter out stop words and produce word
distributions from these messages. These distributions are bucketted
into windows, and then each window is subject to topic analysis and
then we analyze and visualize the results.

@@

We mirrored the repositories and their revisions using software such
as rsync, CVSSuck, \texttt{softChange} and bt2csv.
\texttt{softChange} provided a general schema for storing revisions
and grouped CVS revisions into commits. CVSSuck and rsync mirrored CVS
repositories while bt2csv mirrored web accessible BitKeeper
repositories.

%XXX Smaller?

The documents we are analyzing are the commit log comments i.e., the
comments that are added when revisions to the project are committed.
Per each commit log comment, we count the occurrence of each word in
the message, remove stop words, and then produce word distributions
for each message.  These distributions are then normalized by the size
of the message. After all messages are processed the distributions are
extended to include all words from all of the distributions.

@@

Given all messages, we group the messages into windows. We could use
overlapping windows, but in this paper we use non-overlapping windows
of a set time unit because it simplifies analysis.  Windowing by time
allows for many levels of granularity.  We used one month as the
length of our time windows. While we could use different window
lengths for this study we think that a month is a sizable unit of
development, which is large enough to show shifts in focus, but coarse
enough not to show too much detail.
